---
title: "LogRegEquiv Vignette"
author: "Guy Ashiri-Prossner"
date: "`r format(Sys.Date(), '%d.%m.%Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE
)
```

This package provides tools for assessing the differences between logistic 
regression model-fits across sub-populations. For example, having data with
sub-populations based on gender or native language may or may not require 
separate models.

Consider two sub-populations of Portuguese Language course students, with 
29 covariates and an output variable indicating whether the student has failed 
the course. These sub-populations differ by gender (M/F). These two 
sub-populations provide two logistic regression models, which may or may not 
be equivalent.

The purpose of this vignette is to exemplify to usage of equivalence tests
for different stages in the model: the linear coefficients, the vector of 
per-example log odds ratio and the mean-square error of prediction.
Each method provides a different aspect of possible model equivalence.

The data used is taken from the Student Performance Data[^1]^,^[^2].

[^1]: P. Cortez and A. Silva. Using Data Mining to Predict Secondary School Student Performance. In A. Brito and J. Teixeira Eds., Proceedings of 5th FUture BUsiness TEChnology Conference (FUBUTEC 2008) pp. 5-12, Porto, Portugal, April, 2008, EUROSIS, ISBN 978-9077381-39-7. See also <http://www3.dsi.uminho.pt/pcortez/student.pdf>
[^2]: Data retrieved from [the UC Irvine Machine Learning Repository](https://archive.ics.uci.edu/ml/datasets/student+performance).


```{r setup}
library(LogRegEquiv)
```

# Descriptive Equivalence Testing

The regression coefficient vectors provide us an insight to how a model
describes the phenomenon. In our case, how is the binary variable `final_failure`
affected by the covariates. When two models describe a phenomenon in a similar
manner, descriptive equivalence is achieved.

```{r model construction, echo=TRUE}
formula <- "final_fail ~ ."
female_model <- glm(formula = formula, family = binomial(link = "logit"),
                    data = ptg_stud_f_train)
male_model <- glm(formula = formula, family = binomial(link = "logit"),
                    data = ptg_stud_m_train)
```

Denote $\delta_\beta$ as the maximal tolerated difference per coefficient, in this case
$\delta_\beta=0.1$.

```{r beta equivalence, echo=TRUE}
delta_beta <- 0.1
print(beta_equivalence(model_a = female_model,
                       model_b = male_model,
                       delta = delta_beta,
                       alpha = 0.05))
```

It is possible to assess descriptive equivalence even without creating the models, just by providing the `descriptive_equiv` function a regression formula, two datasets and a `delta` argument:

```{r coef vector equivalence, echo=TRUE}
print(descriptive_equiv(data_a = ptg_stud_f_train,
                              data_b = ptg_stud_m_train, 
                              formula = formula,
                              delta = delta_beta,
                              alpha = 0.05))
```

It is also possible to use different $\delta_\beta$ value for each coefficient. In such case, the `delta` argument should be provided with a vector whose length matches the number of covariates in the model. Mind the intercept and the fact that categorical variables with $k\geq 3$ values are represented in the model by multiple variables. In this example the data has 29 covariates but they are represented by `r length(female_model$coefficients)` covariates in the model:

```{r beta vectorial delta, echo=TRUE}
set.seed(1)
delta_beta_vec <- 0.01 * runif(39)
print(beta_equivalence(model_a = female_model,
                       model_b = male_model,
                       delta = delta_beta_vec,
                       alpha = 0.05))
```


# Individual Predictive Equivalence Testing

Assume we have an existing model $M^A$ based on population $A$. This is our *source*. Given a new *target* population $B$ of size $k$ and model $M^B$, we would like to check whether the source model fits the target population. We do so by comparing the log-odds produced by the models for the *target* population samples. When two models produce equivalent log-odds, individual predictive equivalence is achieved.

To set the indifference level for the individual predictive equivalence $\delta_\theta$, we propose looking 
at the distances between the estimated log odds and the classification threshold. 
Consider $\{\hat{\theta}_1,...,\hat{\theta}_m\}$ to be the $m$ estimated log-odd values for the test population for the original model, meaning that these are predictions for a test set of population A $X_{test}^A$ using model $M_A$. 
In a calibrated logistic regression, the classification of the $i$th subject by the model ($\hat{y}_i$) is 1 when $\hat{\theta}_i >0$ and 0 otherwise. The absolute log-odds $\left| \hat{\theta}_i \right|$ is minimal change to the log-odds that could change to the classification.
We therefore propose setting the threshold $\delta_\theta$ to be a small quantile $r$ (say $r = 0.1$) of the observed distribution of absolute log-odds 
\[\delta_\theta =  \left| \hat{\theta}\right|_{(\lceil r \cdot m \rceil)}. \]

Using $r=0.05$, we get that both models provide equivalent log-odds for the female testing data:

```{r individual predictive equivalence female test, echo=TRUE}
r <- 0.05
print(individual_predictive_equiv(model_a = female_model,
                  model_b = male_model,
                  test_data = ptg_stud_f_test,
                  r = r,
                  alpha = 0.05))
```

On the other hand, using $r=0.025$ we get that both models do not provide equivalent log-odds for the male testing data:

```{r individual predictive equivalence male test, echo=TRUE}
r <- 0.025
print(individual_predictive_equiv(model_a = female_model,
                  model_b = male_model,
                  test_data = ptg_stud_m_test,
                  r = r,
                  alpha = 0.05))
```


# Performance Equivalence Testing

The Brier score is reducing the assessment of the overall model performance into a single figure. Given models $M^A,M^B$ and a test set $X^{test}$ of size $m$, we can assess the equivalence of the Brier scores obtained by the models. The acceptable score degradation $\delta_B>1$ should be selected, we then see the Brier scores as equivalent if $\frac{1}{\delta_B^2}<\frac{BS^B}{BS^A}<\delta_B^2$.
The `dv_index` variable indicates the column number of the dependent variable.

```{r performance equivalence female test, echo=TRUE}
testing_data <- ptg_stud_m_test
print(performance_equiv(model_a = female_model, 
                        model_b = male_model, 
                        test_data = testing_data,
                        dv_index = 30,
                        delta_B = 1.1,
                        alpha = 0.05))


print(performance_equiv(model_a = female_model, 
                        model_b = male_model, 
                        test_data = ptg_stud_f_test,
                        dv_index = 30,
                        delta_B = 1.1,
                        alpha = 0.05))
```
